---
title: "Homework 2"
author: "Adeline Guthrie"
date: "September 11, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(readr)
library(ramify)
library(swirl)
```

## Problem 3

Version control is helpful in the classroom for collaborative coding projects, allowing students to edit code that belongs to a group, still preserving other versions of the code so changes may be checked by other members of the group before finalized.  It also allows the instructor to keep better track of who in each group has been working on the code, possibly to determine if each group member has done their share.


## Problem 4
### Part a
```{r sensory}
#Store URL
url <-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"

#Read in Data, skip first line with just "Operator", treat the second row as the header,
#and fill empty spaces with "NA"
sensory <-read.table(url, header = T, sep = " ", skip = 1, fill = T)

#Shift data values to the right in the rows missing the item number and add the item number
c <- 0
for( i in 1:30){
  if(i%%3 != 1){
    sensory[i,2:6] <- sensory[i, 1:5]
    sensory[i,1] <- c
  }else{
    c <- c + 1
  }
}

#Give each observation its own row
sensory <- gather(sensory, X1, response, -Item)

#Name columns
colnames(sensory) <- c("Item", "Operator", "Response")

#Remove X's infront of operator numbers
sensory$Operator <- parse_number(sensory$Operator)

#Final Dataset:
sensory

#Summarize data
sensory %>%
  group_by(Operator) %>%
  summarise(Mean = mean(Response), Std_dev = sd(Response), Min = min(Response), 
            Median = median(Response), Max = max(Response))

```


### Part b
```{r jumps}
#Store URL
url2 <- "https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"

#Read in data, skip the first row since the labels are messy, fill empty spaces with "NA"
jumps <- read.table(url2, header = F, sep = " ", skip = 1, fill = T)

#Resize the original matrix to give each observation its own row
jumps <- resize(jumps, nrow = 24, ncol =2, byrow = T)

#Remove rows of "NA"
jumps <-na.omit(jumps)

#Name columns
colnames (jumps) <- c("Year", "Long_Jump")

#Convert to tbl_class to use dplyr functions
jumps <- tbl_df(jumps)

#Sort by Year 
jumps <- arrange(jumps, Year)

#Final Dataset:
jumps

#Summarize data
summary(jumps$Long_Jump)

```

### Part c
```{r weight}
#Store URL
url3 <-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"

#Read in data, skip the first row since the labels are messy, fill empty spaces with "NA"
weights <-read.table(url3, header = F, sep = " ", skip = 1, fill = T)

#Resize the original matrix to give each observation its own row
weights <- resize(weights, nrow = 63, ncol =2, byrow = T)

#Name columns
colnames(weights) <- c("Body_Wt", "Brain_Wt")

#Final Dataset:
weights

#Summarize data
summary(weights)

```

### Part d
```{r tomatoes}
#Store URL
url4 <-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"

#Read in data, skip first row and use the second row as header, remove the 
#comment character = "#"
tomatoes <- read.table(url4, header = T, sep = "", skip = 1, comment.char = "")

#Split the data in each of the six cells by ","
I_X10000 <- do.call("rbind", strsplit(toString(tomatoes[1,1]), ","))
I_X20000 <- do.call("rbind", strsplit(toString(tomatoes[1,2]), ","))
I_X30000 <- do.call("rbind", strsplit(toString(tomatoes[1,3]), ","))
P_X10000 <- do.call("rbind", strsplit(toString(tomatoes[2,1]), ","))
P_X20000 <- do.call("rbind", strsplit(toString(tomatoes[2,2]), ","))
P_X30000 <- do.call("rbind", strsplit(toString(tomatoes[2,3]), ","))

#Combine the vectors of split data
I <- cbind(I_X10000, I_X20000, I_X30000)
P <- cbind(P_X10000, P_X20000, P_X30000)

tomatoes <- rbind(I,P)

#Name columns
colnames(tomatoes) <- c("10000_1", "10000_2", "10000_3","20000_1", "20000_2", 
                        "20000_3", "30000_1", "30000_2", "30000_3")

#Convert to tbl_class to use dplyr and tidyr functions
tomatoes <- tbl_df(tomatoes)

#Create column with variety names (I = Ife\#1, P = PusaEarlyDwarf)
tomatoes <- mutate(tomatoes, "Variety" = c("I", "P")) 

#Reorder columns
tomatoes <- select(tomatoes, "Variety", "10000_1", "10000_2", "10000_3","20000_1",
                   "20000_2", "20000_3", "30000_1", "30000_2", "30000_3" )

#Give each observation its own row
tomatoes <- gather(tomatoes, X1, Response, -Variety)

#Fix density labels and remove unnecessary column
tomatoes <- separate(tomatoes, X1, c("Density", "temp"))
tomatoes <- select(tomatoes, Variety, Density, Response)

#Sort by variety and then density
tomatoes <- arrange(tomatoes, Variety, Density)

#Convert Response type to Numeric
tomatoes <- transform(tomatoes, Response = as.numeric(Response))

#Final Dataset:
tomatoes

#Summarize data by variety
tomatoes %>%
  group_by(Variety) %>%
  summarise(Mean = mean(Response), Std_dev = sd(Response), Min = min(Response), 
            Median = median(Response), Max = max(Response))

#Summarize data by density 
tomatoes %>%
  group_by(Density) %>%
  summarise(Mean = mean(Response), Std_dev = sd(Response), Min = min(Response), 
            Median = median(Response), Max = max(Response))


```

## Problem 5

```{r plants}

# Path to data
.datapath <- file.path(path.package('swirl'), 'Courses',
                      'R_Programming_E', 'Looking_at_Data',
                      'plant-data.txt')
# Read in data
plants <- read.csv(.datapath, strip.white=TRUE, na.strings="")

# Remove annoying columns
.cols2rm <- c('Accepted.Symbol', 'Synonym.Symbol')
plants <- plants[, !(names(plants) %in% .cols2rm)]

# Make names pretty
names(plants) <- c('Scientific_Name', 'Duration', 'Active_Growth_Period',
                   'Foliage_Color', 'pH_Min', 'pH_Max',
                   'Precip_Min', 'Precip_Max',
                   'Shade_Tolerance', 'Temp_Min_F')

#Convert to tbl_class to use dplyr functions
plants <- tbl_df(plants)

#Remove rows with NA values
plants <- na.omit(plants)

#Add column "pH_Mid" considering the midpoint of ph_Min and pH_Max
plants <- mutate(plants, 'pH_Mid' = ((pH_Min + pH_Max) / 2))

#Final dataset
plants

#Summarize data by foliage color
plants %>%
  group_by(Foliage_Color) %>%
  summarise(Mean = mean(pH_Mid), Std_dev = sd(pH_Mid), Min = min(pH_Mid), 
            Median = median(pH_Mid), Max = max(pH_Mid))

#Fit linear model
plants.lm <- lm(pH_Mid ~ Foliage_Color, plants)

#Table of coefficients
summary(plants.lm)

#ANOVA results
plants.anova <- aov(pH_Mid ~ Foliage_Color, plants)
summary(plants.anova)

#Boxplot
plot(pH_Mid ~ Foliage_Color, plants)

```